\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-26 paper.qmd
Updated descriptions for section 7.1

@@

<!-- Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>). -->
<!-- ::: -->

In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $b_1$ using UMAP with different (hyper-)parameter settings and tSNE with default (hyper-)parameter setting (@fig-pbmc-mse) were calculated. After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error. Therefore, tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset.

<!--compute absolute error for different parameter choices-->

@@

As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small. Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)). This demonstrates that tSNE accurately captures the data structure of the PBMC3k dataset, which UMAP did not achieve.

>>>>

In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $a_1$ using UMAP, tSNE, PHATE, PaCMAP, and TriMAP with different (hyper-)parameter settings (@fig-pbmc-mse) were calculated. tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset. As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small. Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)).
The visualization of the *model-in-the-data-space* provides evidence for this (@fig-pbmc2-sc). Therefore, tSNE with perplexity $30$ accurately captures the data structure of the PBMC3k dataset, which is better than UMAP.

<!-- As shown in @fig-NLDR-variety (e), there are three well-separated clusters, although the separation between the clusters are small. Additionally, non-linear structures can also be observed within the clusters (@fig-NLDR-variety (e)). This demonstrates that tSNE accurately captures the data structure of the PBMC3k dataset, which UMAP did not achieve. -->

<!-- In order to find a reasonable NLDR representation for the PBMC3k dataset, the MSE for different $b_1$ using UMAP with different (hyper-)parameter settings and tSNE with default (hyper-)parameter setting (@fig-pbmc-mse) were calculated. After analyzing the results, it was found that tSNE with default (hyper-)parameter setting (perplexity: $30$) achieved the lowest error. Therefore, tSNE with a perplexity value set to $30$, the default parameter setting, is considered as a reasonable representation for the PBMC3k dataset. -->

@@

<!-- Model in \gD{}, on the UMAP layout, and three views of the fit in projections from $9\text{-}D$, for the PBMC3k data ($(s_1, \ s_2) = (-0.050, \ -0.041)$, $b = 870 \  (30, \ 29)$, $m = 135$, benchmark value to remove large edges is $0.099$) (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/VqqWuE0Jj6A>). -->
<!-- ::: -->
<!--compute absolute error for different parameter choices-->

@@

### 2024-08-26 paper.tex
Updated descriptions for section 7.1

@@

In order to find a reasonable NLDR representation for the PBMC3k
dataset, the MSE for different \(b_1\) using UMAP with different
(hyper-)parameter settings and tSNE with default (hyper-)parameter
setting (Figure~\ref{fig-pbmc-mse}) were calculated. After analyzing the
results, it was found that tSNE with default (hyper-)parameter setting
(perplexity: \(30\)) achieved the lowest error. Therefore, tSNE with a
perplexity value set to \(30\), the default parameter setting, is
considered as a reasonable representation for the PBMC3k dataset.

@@

As shown in Figure~\ref{fig-NLDR-variety} (e), there are three
well-separated clusters, although the separation between the clusters
are small. Additionally, non-linear structures can also be observed
within the clusters (Figure~\ref{fig-NLDR-variety} (e)). This
demonstrates that tSNE accurately captures the data structure of the
PBMC3k dataset, which UMAP did not achieve.

>>>>

In order to find a reasonable NLDR representation for the PBMC3k
dataset, the MSE for different \(a_1\) using UMAP, tSNE, PHATE, PaCMAP,
and TriMAP with different (hyper-)parameter settings
(Figure~\ref{fig-pbmc-mse}) were calculated. tSNE with a perplexity
value set to \(30\), the default parameter setting, is considered as a
reasonable representation for the PBMC3k dataset. As shown in
Figure~\ref{fig-NLDR-variety} (e), there are three well-separated
clusters, although the separation between the clusters are small.
Additionally, non-linear structures can also be observed within the
clusters (Figure~\ref{fig-NLDR-variety} (e)). The visualization of the
\emph{model-in-the-data-space} provides evidence for this
(Figure~\ref{fig-pbmc2-sc}). Therefore, tSNE with perplexity \(30\)
accurately captures the data structure of the PBMC3k dataset, which is
better than UMAP.

@@

@@
